### 2018-02-06 Cars/paper.tex
paper work

Variational Bayes approximates a distribution $p(\theta | z)$ with a family $q(\theta)$, choosing the member of that family that minimises $KL(q(\theta) || p(\theta | z)$, where
\begin{equation}
\label{KL-div}
KL(q(\theta) || p(\theta | z) = \int_{\theta} q(\theta) \left(\log(q(\theta)) - \log(p(\theta | z))\right) d\theta.
\end{equation}
As $p(\theta | z)$ is typically intractable, VB maximises $\mathcal{L}(q)$ where
\begin{equation}
\label{ELBO}
\mathcal{L}(q) = \int_{\theta} q(\theta) \left(\log(p(\theta, z)) - \log(q(\theta))\right) d\theta
\end{equation}
which is equivalent to minimising (\ref{KL-div}).
This is done with stochastic gradient ascent, taking Monte Carlo estimates of $d\mathcal{L}(q) / d \lambda$ where $\lambda$ is the parameter vector of $q(\theta)$.
There are two Monte Carlo estimators of $d\mathcal{L}(q) / d \lambda$, the Score Estimator, 
\begin{equation}
\label{scoreDeriv}
\frac{d\mathcal{L}(q)}{d \lambda} \approx \sum_{i = 1}^L \frac{d \log(q(\theta_i))}{d \lambda} \left(\log(p(\theta_i, z)) - \log(q(\theta_i)) \right), \mbox{ where } \theta_i \sim q(\theta),
\end{equation}
and the Reparameterised Estimator,
\begin{equation}
\label{rpDeriv}
\frac{d\mathcal{L}(q)}{d \lambda} \approx \sum_{i = 1}^M \frac{d f(\lambda, \epsilon_i)}{d \lambda} \frac{d \log(p(\theta_i, z))}{d \theta_i} + \frac{d J(\lambda, \epsilon_i)}{d \lambda}, \mbox{ where } \epsilon_i \sim r(\epsilon)
\end{equation}
where $r(\epsilon)$ is some distribution without any free parameters (eg. Uniform, Standard Normal), $f(\lambda, \epsilon)$ is some differentiable function that transforms $\epsilon$ to $\theta$, and $J(\lambda, \epsilon)$ is the Jacobian Matrix of this function. (\ref{rpDeriv}) typically has significantly lower variance than (\ref{scoreDeriv}), allowing $M$ to be set lower than $L$ which improves the speed of the algorithm; however the requirement for $f$ to exist restricts the class of approximating distributions $q(\theta)$ that can be used.
Notably, a differentiable $f(\lambda, \epsilon)$ does not exist when $q(\theta)$ is a mixture model and the derivative estimator (\ref{scoreDeriv}) must be used. To reduce $L$ and thus the computational cost of the algorithm, following Gunawan, Tran, and Kohn (2017), Randomised Quasi Monte Carlo (RQMC) is used, where numbers in the unit hypercube are generated according to a Sobol Sequence (Sobol 1967), then these are randomised using the scrambled net method (Matousek 1998), and transformed into a draw of $\theta$ using the inverse-CDF of $q(\theta)$. Using RQMC allows (\ref{scoreDeriv}) to converge to the true derivative at a rate $O(L)$ compared to $O(\sqrt{L})$ in standard Monte Carlo.

@@

\frac{d\mathcal{L}(q)}{d \lambda} \approx \sum_{i = 1}^L \frac{d \log(q(\theta_{N+1} | z_{N+1, 1:T}))}{d \lambda} &\left(\log(q(\theta_{N+1} | z_{N+1, 1:S}) - \log(q(\theta_{N+1} | z_{N+1, 1:T})) \right), \nonumber \\
& \mbox{ where } \theta_i \sim q(\theta_{N+1} | z_{N+1, 1:T}).

@@

\frac{d\mathcal{L}(q)}{d \lambda} \approx \sum_{i = 1}^M \frac{d f(\lambda, \epsilon_i)}{d \lambda} \frac{d \log(q(\theta_{N+1} | z_{N+1, 1:S}))}{d \theta_i} + \frac{d J(\lambda, \epsilon_i)}{d \lambda}, \mbox{ where } \epsilon_i \sim r(\epsilon)

@@

where $f(\lambda, \epsilon)$ transforms a draw of $\epsilon$ into a draw from $q(\theta_{N+1} | z_{N+1, 1:T})$.

@@

>>>>

Variational Bayes (Jordan et al. 1999) approximates a distribution $p(\theta | z)$ with a family $q(\theta)$, choosing the member of that family that minimises $KL(q(\theta) || p(\theta | z)$, where
\begin{equation}
\label{KL-div}
KL(q(\theta) || p(\theta | z) = \int_{\theta} q(\theta) \left(\log(q(\theta)) - \log(p(\theta | z))\right) d\theta.
\end{equation}
As $p(\theta | z)$ is typically intractable, VB maximises $\mathcal{L}(q)$ where
\begin{equation}
\label{ELBO}
\mathcal{L}(q, \lambda) = \int_{\theta} q(\theta) \left(\log(p(\theta, z)) - \log(q(\theta))\right) d\theta
\end{equation}
which is equivalent to minimising (\ref{KL-div}), where $\lambda$ is the parameter vector of $q(\theta)$.
This is done with stochastic gradient ascent, taking Monte Carlo estimates of $\delta\mathcal{L}(q, \lambda) / \delta \lambda$ (Hoffman et al. 2013)
and repeating updates of the form
\begin{equation}
\label{gradientAscent}
\lambda^{(m+1)} = \lambda^{(m)} + \rho^{(m)} \frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda} 
\end{equation}
which is guaranteed to converge to a local maximum (Robbins and Munro, 1951) if $\rho^{(m)}, m = 1, \dots, \infty$ satisfies
\begin{align}
&\sum_{m=1}^{\infty} \rho^{(m)} =  \infty \\
&\sum_{m=1}^{\infty} (\rho^{(m)})^2 <  \infty.
\end{align}
In this paper the sequence $\rho^{(m)}$ is provided by the Adam algorithm of Kingma and Ba (2015).

There are two Monte Carlo estimators of $\delta\mathcal{L}(q, \lambda) / \delta \lambda$, the Score Estimator of Ranganath, Gerrish and Blei (2014), 
\begin{equation}
\label{scoreDeriv}
\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda} \approx \sum_{i = 1}^L \frac{\delta \log(q(\theta_i))}{\delta \lambda} \left(\log(p(\theta_i, z)) - \log(q(\theta_i)) \right), \mbox{ where } \theta_i \sim q(\theta),
\end{equation}
and the Reparameterised Estimator of Kingma Welling (2014),
\begin{equation}
\label{rpDeriv}
\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda} \approx \sum_{i = 1}^M \frac{\delta f(\lambda, \epsilon_i)}{\delta \lambda} \frac{\delta \log(p(\theta, z))}{\delta \theta} \bigg\rvert_{\theta = f(\lambda, \epsilon_i)} + \frac{\delta J(\lambda, \epsilon_i)}{\delta \lambda}, \mbox{ where } \epsilon_i \sim r(\epsilon)
\end{equation}
where $r(\epsilon)$ is some distribution without any free parameters (eg. Uniform, Standard Normal), $f(\lambda, \epsilon)$ is some differentiable function that transforms $\epsilon$ to $\theta$, and $J(\lambda, \epsilon)$ is the Jacobian Matrix of this function. (\ref{rpDeriv}) typically has significantly lower variance than (\ref{scoreDeriv}), allowing $M$ to be set lower than $L$ which improves the speed of the algorithm; however the requirement for $f$ to exist restricts the class of approximating distributions $q(\theta)$ that can be used.
Notably, a differentiable $f(\lambda, \epsilon)$ does not exist when $q(\theta)$ is a mixture model and the derivative estimator (\ref{scoreDeriv}) must be used. To offset the increased variance and allow a reduction in $L$, following Gunawan, Tran, and Kohn (2017), Randomised Quasi Monte Carlo (RQMC) is used, where numbers in the unit hypercube are generated according to a Sobol Sequence (Sobol 1967), then these are randomised using the scrambled net method (Matousek 1998), and transformed into a draw of $\theta$ using the inverse-CDF of $q(\theta)$. Using RQMC allows (\ref{scoreDeriv}) to converge to the true derivative at a rate $O(L)$ compared to $O(\sqrt{L})$ in standard Monte Carlo.

@@

\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda} \approx \sum_{i = 1}^L \frac{\delta \log(q(\theta_{N+1, i} | z_{N+1, 1:T}))}{\delta \lambda} &\left(\log(q(\theta_{N+1, i} | z_{N+1, 1:S}) - \log(q(\theta_{N+1} | z_{N+1, 1:T})) \right), \nonumber \\
& \mbox{ where } \theta_{N+1, i} \sim q(\theta_{N+1} | z_{N+1, 1:T}).

@@

\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda_T} \approx \sum_{i = 1}^M \frac{\delta f(\lambda_T, \epsilon_i)}{\delta \lambda} \frac{\delta \log(q(\theta_{N+1} | z_{N+1, 1:S}))}{\delta \theta_{N+1}} \bigg\rvert_{\theta_{N+1} = f(\lambda_T, \epsilon_i)} + \frac{\delta J(\lambda_T, \epsilon_i)}{\delta \lambda_T}, \mbox{ where } \epsilon_i \sim r(\epsilon)

@@

where $\lambda_T$ denotes the parameter vector of $q(\theta_{N+1} | z_{N+1, 1:T})$.

@@

There are six naive forecast models with constant estimates of the future values of $v_t$ and $\delta_t$ for $t = T+1, \dots, T+30$. $v_t$ is estimated by either the current velocity, $v_T$ or the average of the previous ten lags of $v$, $1/10 \sum_{s = T - 9}^T v_s$. $\delta_t$ is estimated by one of the current steering angle, $\delta_T$, the average of the previous ten lags of $\delta$, $1/10 \sum_{s = T - 9}^T \delta_s$ or by $\pi/2$, i.e. the car is moving directly ahead. Table \ref{tableNaive} classifies each naive model based on the combination of velocity and angle estimates.
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|}
\hline
& $\hat{\delta}_t = \delta_T$ & $\hat{\delta}_t = 1/10 \sum_{s=T-9}^T \delta_s$ & $\hat{\delta}_t = \pi/2$ \\
\hline
$\hat{v}_t = v_T$ & Naive 1 & Naive 2 & Naive 3 \\
$\hat{v}_t = 1/10\sum_{s=T-9}^T$ & Naive 4 & Naive 5 & Naive 6\\
\hline
\end{tabular}
\end{center}
\label{tableNaive}
\caption{Classification of naive forecast models by future velocity and steering angle estimators}
\end{table}
